\textbf{AprilTrack: A Particle-Filter-based AprilTag Tracker}

### 2017-10-09 paper/paper.tex
Fixed title
Made typo fixes to paper

\title{\textbf{AprilTrack: A Particle-Filter based AprilTag Tracker}}
\author{Daniel Pfrommer}
\date{}

@@

 The majority of previous literature in fiducial detection has been focused single-frame fiducial detection \citep{RUNETag, BlurResilient, AprilTag}. A common assumption in these algorithms is that the image being processed shows little motion blur, an expectation which breaks down in empirical situations. For instance, the widely-used AprilTag marker \citep{AprilTag} and its successor, the AprilTag 2 \citep{AprilTag_2}, rely on strong gradients between the tag's background and the surrounding white border to detect the tag. As shown in figure \ref{fig:blurred_tag}, oftentimes such algorithms fail to detect tags that have been blurred, either missing them entirely or mistakenly discarding them when eliminating false positives through some encoded fiducial ID (for AprilTags this consists of a 6x6 binary grid in the center of the tag which is used to verify the validity of the detection).

@@

Previously, there has been interest in designing blur resilient fiducials. Meghshyam et al. \citep{BlurResilient} have developed ringed markers to ensure that a region of the tag will always be perpendicular to the blur direction. However, this approach, while resistent to linear blur, does not take into account more complex blurs (e.g resulting from the tag pivoting around the vertical axis). In addition, the increased resilience to blur comes at the cost of localization accuracy as only the center point of the ringed tag can be reliably determined \citep{BlurResilient}, whereas with other fiducials, such as the AprilTag \citep{AprilTag}, full pose estimation in 6 degrees of freedom (DOF) can be done based on just a single tag. Another significant disadvantage of the ringed marker from Meghshyam et al. is the limited number of tag ID's, as their proposed encoding method allows for a maximum of 4 distinct tags \citep{BlurResilient}.

Unlike previous attempts to engineer more resilient tags, the algorithm proposed here exploits the temporal coherence between successive frames in a video sequence to track a tag's location between successful detections. As many high-blur scenarios involve a continuous stream of images taken at equidistant time intervals (e.g a quadroter-mounted moving camera viewing tags or a stationary camera tracking an AprilTag-labeled moving object), we propose to use knowledge of a fiducial's prior location to augment the detection algorithm. As such, this paper presents a novel particle-filter-based tracking algorithm (AprilTrack) capable of tracking AprilTags in 6 DOF with both joint-tag motion and camera ego-motion.

Previous work involving the tracking of blurred objects includes the Blur-driven Tracker (BLUT) framework proposed by Wu et al. \citep{BLUT} for tracking motion-blurred targets. BLUT, a particle-filter based tracker, works on the observation that blur strength and direction can be used estimate an object's velocity and provide valuable information for successfully tracking a blurred object. While capable of reliably tracking blurred objects, BLUT suffers from the disadvantage that it can do so only in two dimensions (i.e it is assumed that the object always faces the camera) and so cannot track a sequence of involving arbitrary tag rotations and translations.

The novel method proposed in this paper augments the AprilTag detection algorithm originally designed by Olson \citep{AprilTag} and improved by Wang et al \citep{AprilTag_2} with a particle-filter based algorithm to reliably track an AprilTag through an arbitrary sequence of images, providing valuable fiducial-based position information when existing methods would be unable to do so. In addition, the proposed algorithm is extended to incorporate the joint motion of an arbitrary number of tags to increase the accuracy of the tracked orientation and position.

@@

\textbf{Particle Filters.}\footnote{A more in-depth discussion of particle filters can found in \citep{ParticleNotes}. Much of the notation used in \citep{ParticleNotes} is similar to that used here.} In this regard, particle filters attempt to approximate the posterior distribution at time $t-1$, $p(\vec{x}_{0:t-1}|\vec{z}_{1:t-1})$, using Monte-Carlo methods. A particle filter consists of a set of $N$ samples $\{ (\vec{x}^i_{t-1}, w^i_{t-1}) \}_{i=1}^N$ drawn from the proposal distribution $q(\vec{x}^i)$ and weighted so as to take into account any discrepancy between $q(\vec{x})$ and $p(\vec{x})$ in that $w^i_t \propto p(\vec{x}^i_{0:t}|\vec{z}_{1:t})/q(\vec{x}^i_{0:t}|\vec{z}_{1:t})$.

@@

As shown in figure \ref{fig:error}, the pixel corner error hovers at around 10 pixels for most of the sequence. The tracker loses the tag momentarily around frames 30 and 200 but is able to recover rapidly without any manual intervention. Much of the error observed throughout the sequence is the result of disagreement on the boundary of a blurred tag rather than a systematic issue with the tracker itself.

@@

We have presented the AprilTrack algorithm, which augments standard AprilTag detection algorithms with a tracker capable of tracking fiducials in 6 DOF. Experimental results show that the tracker is capable of accurately tracking both single tags and consolidating information from multiple tags in ego-motion and joint tag-motion scenarios. In future research we plan to investigate other possible error metrics for patch comparison and look to deploy AprilTrack in realtime scenarios.

>>>>

\title{\textbf{AprilTrack: A Particle-Filter-based AprilTag Tracker}}
\author{Daniel Pfrommer}
\date{}

@@

 The majority of previous literature in fiducial detection has been focused on single-frame fiducial detection \citep{RUNETag, BlurResilient, AprilTag}. A common assumption in these algorithms is that the image being processed shows little motion blur, an expectation which breaks down in practice. For instance, the widely-used AprilTag marker \citep{AprilTag} and its successor, the AprilTag 2 \citep{AprilTag_2}, rely on strong gradients between the tag's background and the surrounding white border to detect the tag. As shown in figure \ref{fig:blurred_tag}, oftentimes such algorithms fail to detect tags that have been blurred, either missing them entirely or mistakenly discarding them when eliminating false positives through some encoded fiducial ID (for AprilTags this consists of a 6x6 binary grid in the center of the tag which is used to verify the validity of the detection).

@@

Previously, there has been interest in designing blur resilient fiducials. Meghshyam et al. \citep{BlurResilient} have developed ringed markers to ensure that a region of the tag will always be perpendicular to the blur direction. However, this approach, while resistent to linear blur, does not take into account more complex blurs (e.g resulting from the tag pivoting around the vertical axis). In addition, the increased resilience to blur comes at the cost of localization accuracy as only the center point of the ringed tag can be reliably determined \citep{BlurResilient}, whereas with other fiducials, such as the AprilTag \citep{AprilTag}, full pose estimation in 6 degrees of freedom (DOF) can be done based on just a single tag. Another significant disadvantage of the ringed marker of Meghshyam et al. is the limited number of tag ID's, as their proposed encoding method allows for a maximum of 4 distinct tags \citep{BlurResilient}.

Unlike previous attempts to engineer more resilient tags, the algorithm proposed here exploits the temporal coherence between successive frames in a video sequence to track a tag's location between successful detections. As many high-blur scenarios involve a continuous stream of images taken at equidistant time intervals (e.g a quadroter-mounted moving camera viewing tags or a stationary camera tracking an AprilTag-labeled moving object), we propose to use knowledge of a fiducial's prior location to augment the detection algorithm. As such, this paper presents a novel particle-filter-based tracking algorithm (AprilTrack) capable of tracking AprilTags in 6 DOF with both joint tag motion and camera ego-motion.

Previous work involving the tracking of blurred objects includes the Blur-driven Tracker (BLUT) framework proposed by Wu et al. \citep{BLUT} for tracking motion-blurred targets. BLUT, a particle-filter based tracker, works on the observation that blur strength and direction can be used to estimate an object's velocity and provide valuable information for successfully tracking a blurred object. While capable of reliably tracking blurred objects, BLUT suffers from the disadvantage that it can do so only in two dimensions (i.e it is assumed that the object always faces the camera) and so cannot track a sequence involving arbitrary tag rotations and translations.

The novel method proposed in this paper augments the AprilTag detection algorithm originally designed by Olson \citep{AprilTag} and improved by Wang et al \citep{AprilTag_2} with a particle-filter based algorithm to reliably track an AprilTag through an arbitrary sequence of images, providing fiducial-based position information when existing methods fail. In addition, the proposed algorithm is extended to incorporate the joint motion of an arbitrary number of tags to increase the accuracy of the tracked orientation and position.

@@

\textbf{Particle Filters.}\footnote{A more in-depth discussion of particle filters can found in \citep{ParticleNotes}. Much of the notation used in \citep{ParticleNotes} is similar to that used here.} Particle filters attempt to approximate the posterior distribution at time $t-1$, $p(\vec{x}_{0:t-1}|\vec{z}_{1:t-1})$, using Monte-Carlo methods. A particle filter consists of a set of $N$ samples $\{ (\vec{x}^i_{t-1}, w^i_{t-1}) \}_{i=1}^N$ drawn from the proposal distribution $q(\vec{x}^i)$ and weighted so as to take into account any discrepancy between $q(\vec{x})$ and $p(\vec{x})$ in that $w^i_t \propto p(\vec{x}^i_{0:t}|\vec{z}_{1:t})/q(\vec{x}^i_{0:t}|\vec{z}_{1:t})$.

@@

As shown in figure \ref{fig:error}, the pixel corner error is about 10 pixels for most of the sequence. The tracker loses the tag momentarily around frames 30 and 200 but is able to recover rapidly without any manual intervention. Much of the error observed throughout the sequence is the result of disagreement on the boundary of a blurred tag rather than a systematic issue with the tracker itself.

@@

We have presented the AprilTrack algorithm, which augments standard AprilTag detection algorithms with a tracker capable of tracking fiducials in 6 DOF. Experimental results show that the tracker can track both single tags and consolidating information from multiple tags in ego-motion and joint tag-motion scenarios. In future research we plan to investigate other possible error metrics for patch comparison and look to deploy AprilTrack in realtime scenarios.
